---
title: "Package refele"
author: "Benjamin Louis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(refele)
```

## Summarise hourly climatic data towards daily climatic data

Climatic data can be obtained at hourly intervals but need to be used at daily scale. Function `summarise_daily` allows to summarise hourly climatic data towards daily climatic data either by calculating a cumulative sum of the data aver each day (e.g. for precipitation data) or by calculating a daily mean value of the data (e.g. for temperature data).

You can see how ti use the function with :

```{r eval = FALSE}
?sumarise_daily
```

Here is an example with 3-hours interval climatic dataset :

```{r}
library(dplyr, warn.conflicts = FALSE)
library(lubridate, warn.conflicts = FALSE)
# Convert date time column to date time class
hourlydat <- meteoClermont %>%
  mutate(date = ymd_hms(date))
# Run the function
dailydat <- summarise_daily(.data = hourlydat, .timevar = date, .cumul = precipitation , 
                .average = wind_dir:pres)
```

`summarise_daily` accept a dataframe (`.data`) with a column corresponding to the date and time of data measurement (`.timevar`). If the classe of `.timevar` is not a R Date-Time class (`POSIXct`), the function will try to parse it but you will get a warnings that you should do it on your own before, to be sure that the good format is used. In this example, function `lubridate::ymd_hms` is used in order to do it. 

Argument `.cumul` is used to give the name of columns that need to be summed and `.average` is used to the column that need to be averaged. Here, all columns betweens `wind_dir` and `pres` are averaged and only the column `precipitation` is summed.

We can see the graph of wind direction measured at hourly intervals (black dots) and its daily mean (blue line) :

```{r warning = FALSE, out.width = '95%', fig.width = 7, fig.asp = 1, dpi = 300, fig.align = 'center'}
library(ggplot2)
dat <- left_join(
  select(hourlydat, date, wind_dir_hourly = wind_dir) %>% 
    mutate(the_days = floor_date(date, unit = "day")),
  select(dailydat, the_days, wind_dir_daily = wind_dir),
  by = "the_days"
)
ggplot(dat) +
  geom_point(aes(x = date, y = wind_dir_hourly), shape = 20, colour = "black") +
  geom_line(aes(x = date, y = wind_dir_daily), colour = "blue") +
  labs(x = "Date", y = "Wind direction")
```


